---
title: "Lab04 - F tests for ANOVA"
output: pdf_document
---

## Goals

The goal in this lab is to practice F tests for ANOVA

## Loading packages

Here are some packages with functionality you may need for this lab.  Run this code chunk now.

```{r}
library(readr)
library(ggplot2)
library(gridExtra)
library(mosaic)
library(dplyr)

options("pillar.sigfig" = 10) # print 10 significant digits in summarize output
```

A study was conducted to examine the effectiveness of four different hand-washing methods for eliminating bacteria

```{r}
soap <- read_csv("http://www.evanlray.com/data/sdm4/Bacterial_Soap.csv")
names(soap) <- c("bacterial_counts", "method")

soap %>%
  count(method)
```

#### 1. Make an appropriate plot of the data.

```{r}
ggplot(data = soap, mapping = aes(x = method, y = bacterial_counts)) +
  geom_boxplot()

ggplot(data = soap, mapping = aes(x = bacterial_counts, color = method)) +
  geom_density()

ggplot(data = soap, mapping = aes(x = bacterial_counts)) +
  geom_histogram(bins = 20) +
  facet_wrap( ~ method, ncol = 1)
```

#### 2. Conduct a test of the claim that all four methods are equally effective.

Define the parameters

$\mu_1$ = mean bacterial count when using alcohol spray

$\mu_2$ = mean bacterial count when using antibacterial soap

$\mu_3$ = mean bacterial count when using soap

$\mu_4$ = mean bacterial count when using water

Our hypotheses are:

$H_0: \mu_1 = \mu_2 = \mu_3 = \mu_4$  The average bacterial count is the same for all four handwashing methods.

$H_A:$ Not all of $\mu_1$, $\mu_2$, $\mu_3$, and $\mu_4$ are equal to teach other.  The average bacterial count is different for at least one of the handwashing methods.

```{r}
fit_full <- lm(bacterial_counts ~ method, data = soap)
anova(fit_full)
```

The p-value for this test is 0.001111.  The data provide very strong evidence that the mean bacterial counts are not equal for the four handwashing methods.

#### 3. Conduct a test of the claim that the "Antibacterial Soap", "Soap", and "Water" methods are equally effective.

```{r}
soap <- soap %>%
  mutate(
    not_alcohol_spray = ifelse(method %in% c("Antibacterial Soap", "Soap", "Water"), "Not Alcohol", "Alcohol")
  )
fit_reduced <- lm(bacterial_counts ~ not_alcohol_spray, data = soap)
anova(fit_reduced, fit_full)
```

In terms of the parameters defined in the answer to part 2, we are testing the following hypotheses:

$H_0: \mu_2 = \mu_3 = \mu_4$  The mean bacterial count is the same for each of the antibacterial soap, soap, and water methods.

$H_A:$ At least one of $\mu_2$, $\mu_3$, and $\mu_4$ is different from the other two.  The mean bacterial count is not the same for all three of these methods.

The p-value for the test is 0.436.  The data do not provide strong evidence that there is a difference in mean bacterial counts.

#### 4. (Optional) If you have extra time and want a refresher, calculate the mean and standard deviation of observations in each group

```{r}
soap %>%
  group_by(method) %>%
  summarize(
    mean_bacteria = mean(bacterial_counts),
    sd_bacteria = sd(bacterial_counts)
  )
```
